General constraints on magic distillation protocols from statistical mechanics

### 2021-05-28 magic_paper.tex
punctuation and typos

\begin{document}
% Labels and Refs (consistency, thm vs lem vs prop)
% Figs check(fig.0, move to app, consistency)

% Notation consistency (R is rate not bound, ket vs |> etc, \ln{} -> \ln ,??, italics)
% UK vs US english
% Spacing (figs, sections, paragraphs, eq. numbering, \noindent, QEDs)

@@

In this work, we develop a framework to analyse magic state distillation protocols where explicit physics of the system is incorporated.  These distillation bounds are obtained by considering how a given magic state protocol would behave if it were instead applied to a \emph{stabilizer} state that is distinguished by the physics of the system -- for example an equilibrium state at a characteristic temperature, or the maximally mixed state. The magic distillation bounds can then be expressed in terms of this distinguished degrees of freedom (e.g. free energy changes in the case of temperature, or non-unitality in the case of the maximally mixed state). Such bounds are therefore of interest in assessing how physical features such as temperature, noise-biases or fixed-point structure associated with restricted gate-sets constrain distillation protocols, and so of relevance to existing work towards fault-tolerant quantum computing~\cite{Aliferis_2008, Stephens_2013, Li_2015, Babbush_2018, Tuckett_2019, Guillaud_2019, Fowler_2019}.

The approach we take is most closely aligned with resource theories of magic, however it differs in a few ways. Firstly, we obtain distillation upper bounds without the use of monotones, and secondly our distillation bounds are hardware dependent. We approach this problem using tools from statistical mechanics and recent work from the overlap of thermodynamics and quantum information theory~\cite{cit:janzing, cit:horodecki2, Brandao_2015, Vinjanampathy_2016, Goold_2016, cit:lostaglio, cit:gour}. In particular, we shall view \emph{magic as a form of free energy}, relative to the set of stabilizer states. From this perspective the convex set of stabilizer states behaves as the set of ``equilibrium'' states, and magic corresponds to a form of non-equilibrium free energy that is strictly non-classical.

@@

Consider a $d$--dimensional quantum system with Hilbert space $\H_d$, and let $\{ |0\>, |1\>, \dots , |d-1\>\}$ denote the standard computational basis, defined over $\mathbb{Z}_d = \{ 0, 1, \dots,d-1 \}$. On this space, generalised Pauli matrices $X, Z$ can be defined by their respective roles as shift and phase operators, acting on the basis states as follows,
\begin{align}
    X \ket{k} &= \ket{k + 1} \label{eq:xpauli}\\
	Z \ket{k} &= \omega^k \ket{k}. \label{eq:zpauli}
\end{align}
Here $\omega \coloneqq e^{2\pi i/d}$ is the $d$-th root of unity and addition is taken modulo $d$. From these we can construct a phase space $\cal{P}_d = \mathbb{Z}_d \times \mathbb{Z}_d$ that provides a complete representation of the quantum system. Given a point $\bmx \coloneqq (x, p)$ we define a displacement operator, 
\begin{equation}\label{eq:ddef}
    D_{\bmx} \coloneqq \tau^{x p} X^{x} Z^{p},\ 
\end{equation}
where the phase factor $\tau \coloneqq -\omega^{1/2}$ ensures unitarity. We assume going forward that $d$ is an odd prime, however the case $d=2$ can also be handled, but with some additional technical caveats~\cite{Appleby_2005}. For a composite system with composite dimension $d = d_1 \dots d_n$ we can decompose the Hilbert space as $\cal{H}_d = \cal{H}_{d_1} \otimes \dots \otimes \cal{H}_{d_n}$, and then define displacement operators as
\begin{equation}\label{eq:composited}
    D_{\bmx} \coloneqq D_{(x_1, p_1)} \otimes \dots \otimes D_{(x_n, p_n)},
\end{equation}
where now we have
\begin{align*}
	\bmx \coloneqq (x_1, p_1, x_2, p_2, \dots, x_n, p_n) \in \cal{P}_{d_1} \times \dots \times \cal{P}_{d_n} \eqqcolon  \cal{P}_d,
\end{align*}
to denote the phase space point for the composite system. 
For simplicity going forward, we assume $n$ copies of a $d$--dimensional system $d_1=d_2 = \cdots = d$, and therefore, we have that $\x \in \mathbb{Z}_d^{2n}$.

@@

\begin{figure}[t]
    \centering
    \subfigure[][]{%
    \label{fig:maxmix}%
    \includegraphics[height=2cm]{figs/maxmixed.pdf}
    %\caption{Maximally mixed state $\frac{1}{3}\id$}%
    }\hspace{8pt}%
    \subfigure[][]{%
    \label{fig:zero}%
    \includegraphics[height=2cm]{figs/zerostate.pdf}
    %\caption{Zero state $\ketbra{0}{0}$}%
    }\\
    \subfigure[][]{%
    \label{fig:bound}%
    \includegraphics[height=2cm]{figs/boundstate.pdf}
    %\caption{Bound state}%
    }\hspace{8pt}%
    \subfigure[][]{%
    \label{fig:Strange}%
    \includegraphics[height=2cm]{figs/Strangestate.pdf}
    %\caption{Strange state $\ketbra{S}{S}$}%
    }
    \caption{\textbf{Qutrit Wigner distributions of varying magic.} 
    \subref{fig:maxmix} Maximally mixed state $\frac{1}{3}\id$; \subref{fig:zero} Stabilizer zero state $\ketbra{0}{0}$; \subref{fig:bound} A non-stabiliser Wigner-positive state; \subref{fig:Strange} Magic Strange state $\ket{{\rm{S}}} = \frac{1}{\sqrt{2}}(\ket{1} - \ket{2})$, coined in~\cite{cit:veitch2}.
    }%
    \label{fig:wstate_examples}
\end{figure}

@@

Any magic state protocol will correspond to a quantum channel $\E$, and admits a Wigner representation $W_{\E}( \y | \x)$ that can be viewed as a transition matrix mapping phase space points $\x \rightarrow \y$. The representation obeys the relation 
\begin{equation}
	W_{\E(\rho)} (\y) = \sum_{\x \in \P_d} W_{\E}( \y | \x) W_\rho(\x),
\end{equation}
for any $\E$ and $\rho$. Now if any magic protocol will send free states to free states, and so $\E$ sends all positively represented quantum state to other positively represented quantum states. In such cases, it can be shown~\cite{Wang_2019} that $W_{\E}(\y |\x)$ must form a stochastic matrix. In particular, all Clifford operations correspond to stochastic matrices in this Wigner representation.

@@

Our analysis applies majorization to magic states at the level of the associated Wigner distributions. However, these are in general quasi-probability distributions and so it is important to check how majorization is computed for these cases and what differences quasi-distributions bring over genuine probability distributions.

@@

However, if $\w$ is a \emph{quasi-probability} distribution with negative values, and $\r$ a regular probability distribution things are different. Now the Lorenz curve is no longer monotone increasing, but is a concave function that breaks through the $L_{\w|\r}(x) = 1$ barrier at an interior point and attains some non-trivial maximum $L_\star$ above the value $1$, before decreasing monotonically to $L_{\w|\r}(x)= 1$ at the end-point $x=1$. See Figure \ref{fig:lcs} for examples of non-monotonic Lorenz curves for quasi-distributions. Within quantum theory, the breaking of this barrier is associated with the degree of non-classicality. Since majorization provides a measure of order/disorder this is consistent with the intuition that negativity, for example within the context of quantum optics or quantum computing, can provide a greater form of order than is possible within classical theory.

\begin{figure}
    \centering
    \includegraphics[scale=0.35]{figs/lc_strange.pdf}
    \caption{\textbf{A `heretical' family of Lorenz curves.} Traditionally, Lorenz curves for distributions are monotone increasing cumulant functions that reach a maximum value of $1$. In contrast, Lorenz curves for magic states break through the value of $L(x)=1$ due to the presence of negativity in the associated quasi-probability distributions. The above family of curves correspond to multiple copies of noisy Strange states $\rho\coloneqq\rho_{\rm{S}}(\epsilon)^{\otimes n}$ (\cref{eq:noisysn}) for $n=2,4,6$ within a particular stabilizer fragment ($\sigma = \id/3$). Solid lines represent pure Strange states, while dashed lines represent $\epsilon$-noisy Strange states with depolarising noise $\epsilon = 0.1$.
    }
    \label{fig:lcs}
\end{figure}

@@

 In Section \ref{sec:stab} we shall focus on the energetic costs of a protocol, that are defined due to the specific Hamiltonian that the system possesses. However, before that, in this section we encode physical constraints at a more abstract level in terms of the fixed-point structure of  the actual operations that are accessible to us. We show that any magic theory can always be decomposed in terms of `sub-theories' with a particular fixed-point structure. This section therefore exploits $\mathbf{d}$--majorization, while in Section \ref{sec:stab} we drop the requirement of fixed-points and instead use the full relative majorization relation.

 We begin by defining the following sub-theory of any magic theory $\R$.
\begin{definition}\label{def:sigmafrag}
   Given a theory of magic $\R = (\F, \O)$, the \emph{$\sigma$--fragment of $\R$} is the sub-theory $\R_\sigma = (\F, \O_\sigma)$, in which the free operations are 
   \begin{equation}
        \O_\sigma \coloneqq \{ \E \in \O: \E(\sigma) = \sigma \},
    \end{equation}
namely those operations that leave the state $\sigma \in \F$ invariant.
\end{definition}
This notion provides a simple way to break up any theory into smaller, more manageable parts that are destinguished by additional physical constraints, illustrated in~\cref{fig:zoo}. Moreover, the union over all fragments returns the parent theory, so we are not discarding any information by breaking up the theory in this way. We make this statement precise as follows.
\begin{theorem}\label{thm:frag}
    Let $\R = (\F, \O)$ be a theory of magic.
Then a transformation $\rho \rightarrow \tau$ is possible in $\R$ if and only if the transformation $\rho \rightarrow \tau$ is possible in at least one $\sigma$--fragment of $\R$.
\end{theorem}
The proof of this is straightforward.
\begin{proof}
   Suppose the interconversion is possible in a $\sigma$--fragment via some $\E \in \O_\sigma$. But since $\O_\sigma \subseteq \O$ it is also possible in $\R$. Conversely, suppose the transformation is possible in $\R$ via some $\E$ in $\O$. The free states $\F$ are a closed, bounded set and moreover the image of $\F$ under the map $\E$ is in $\F$. By the Brouwer Fixed-Point theorem~\cite{cit:brouwer} this mapping must therefore have a fixed point $\sigma \in \F$. Therefore $\E \in \O_\sigma$ and so the interconversion is possible in this fragment.
\end{proof}

@@

If we consider some $\sigma \in \F$, and its corresponding fragment $\R_\sigma$, then we are typically interested in the ability to transform many copies of some magic state $\rho$ into more pure forms of magic. The state $\rho$ is assumed to have negativity in the Wigner representation, so $W_\rho(\x) < 0$ for some regions of $\x \in \P_d$. We again note that we restrict to odd-dimensional quantum systems, for simplicity. In contrast, the state $\sigma$ will have a Wigner distribution $W_\sigma(\x)$ that forms a genuine probability distribution. On the technical assumption that $\sigma$ is full rank, we also have that $W_\sigma(\x) > 0$ for all $\x \in \P_d$. Any non-full rank states $\sigma$ can be handled as a limiting case in which we inject an infinitesimal fraction of depolarising noise $\epsilon (\id/d)$.

@@

If we have a magic state $\rho$ that has negativity in its Wigner representation, then, as discussed, its Lorenz curve over-shoots $1$ and reaches a non-trivial maximum $L_\star$ that depends on the particular state. There is an exact relation between $L_\star$ and sum-negativity or mana, which is provided by the following result. 
\begin{theorem}\label{lem:lcmax}
	Given a magic state $\rho$, the maximum $L_\star$ of its Lorenz curve $\lc{\rho}{\sigma}(x)$ is independent of the $\sigma$--fragment and equal to $1+\sn{\rho}$. Moreover, the majorization constraint is stronger than mana in every fragment.
\end{theorem}
The proof of this is given in~\cref{app:major}. Using the Lorenz curve perspective, it is simple to construct other monotones. For example, within a given fragment, the area above the horizontal line $y = 1$ is another magic monotone.
\begin{lemma}
Given a magic state $\rho$ and a free state $\sigma$, let $\A_\sigma(\rho)$ be the area of the region $\{(x, y): 1 \leq y \leq L_{\rho | \sigma}(x)\}$. Then $\A_\sigma(\rho)$ is a magic monotone for $\R_\sigma$.
\end{lemma}
\begin{proof}
Consider the transformation $\rho_1 \rightarrow \rho_2$ within $\R_\sigma$. We have that $L_{\rho_1|\sigma}(x)$ is never below the curve $L_{\rho_2|\sigma}(x)$, and therefore the region above $L=1$ for $\rho_2$ is a subset of the corresponding region for $\rho_1$. Thus, $\A_\sigma(\rho_1) \geq \A_\sigma(\rho_2)$, so $\A_\sigma$ is a magic monotone in $\R_\sigma$.
\end{proof}
In contrast to mana, the area monotone is \emph{specific to the fragment}, and its value will vary as we change $\sigma$. Therefore its monotonicity is also including the physics of the fragment and provides a means to analyse magic distillation only under free operations that leave $\sigma$ invariant.

@@

For our distillation protocols we need to choose a canonical pure magic state as a target. For this we shall choose the Strange state $|S\>$ given as
\begin{equation}
|S\> \coloneqq \frac{1}{\sqrt{2}} (|1\> + |2\>),
\end{equation}
in the computational basis. This qutrit magic state is exceptionally symmetric and simple which benefits our analysis. Its distribution $W_S(\x)$ has a single negative value of $-1/3$ at $\x =(0,0)$ and the positive value $1/6$ at all other points. We can define the $\epsilon$--noisy Strange state as
\begin{equation}\label{eq:noisysn}
    \rho_{\rm{S}}(\epsilon)\coloneqq (1 - \epsilon) \ket{\rm{S}}\bra{\rm{S}} + \epsilon \frac{1}{3}\id,
\end{equation}
where $\epsilon$ is the depolarization noise parameter. Moreover, utilising the symmetries of the Strange state allow for any magic state $\rho$ to be processed via Clifford operations~\cite{cit:prakash,cit:prakash2} and put into this canonical form for some $\epsilon \ge 0$.

The Wigner distribution of the single-copy, $\epsilon$--noisy Strange state  is given by
\begin{equation}
	\W[\bmx]{\rho_{\rm{S}}(\epsilon)} = (1-\epsilon)\W[\bmx]{\ketbra{\rm{S}}} + \epsilon\W[\bmx]{\frac{1}{3}\id}.
\end{equation}
Since the Wigner distribution $W_{\I/3}(x)$ is the uniform probability distribution over the phase space, we get that $W_{\rho_S(\epsilon) } (\x)$ has a single negative component
\begin{equation}
	- v(\epsilon) \coloneqq - \left( \frac{1}{3} -\frac{4}{9}\epsilon \right)
\end{equation} 
located at the origin $\bmx = \bmo$ and positive components
\begin{equation}
	u(\epsilon) \coloneqq \frac{1}{6} -\frac{1}{18}\epsilon
\end{equation}
at the 8 phase space points $\x \ne (0,0)$. We assume that $\epsilon < 3/4$ to ensure the presence of negativity in the Wigner distribution.

@@

In the next section we leave fragments behind, and instead use the full structure of relative majorization to obtain bounds that incorporate energetic aspects of a magic state protocol.

@@

\subsection{Free energy and sub-linear correlations in the thermodynamic limit}
To simplify our analysis we will make the following physical assumption. We assume that in the asymptotic/thermodynamic limit $n,m \rightarrow \infty$ that correlations generated between subsystems are negligible. This implies that $d\mu(x)$ is peaked on a particular state $\tau'$, and so $\E(\tau) = \tau^{\prime \otimes m}$. Physically, this is a natural scenario -- for example, in the context of traditional thermodynamics, it states that the output system is well-described by intrinsic variables that are encoded in $\tau'$, and correlations are sub-linear in $m$. This assumption allows us to associate a free energy per qudit of the output state, which would be more complex if non-trivial correlations were present.

@@

\subsection{Temperature-dependent distillation bounds}
The preceding discussion gives a means to specify energetic details of a protocol through how much it would disturb some reference equilibrium state. When combined with relative majorization, it also allows us to derive magic distillation bounds that combine both the computational measures $\epsilon,\epsilon'$, with terms that depend on the Hamiltonian of the physical system. As discussed in section~\nick{REF}, from the perspective of majorization the presence of magic can be viewed as a form of non-classical free energy and so it is also of interest to make an explicit link between the amount of magic in a non-equilibrium state $\rho_S(\epsilon)$ and the actual free energy of an equilibrium state $\tau(\beta)$.

@@

Then the magic distillation rate $R = m/n$ for the protocol is bounded by the expression
\begin{equation}\label{eq:rate_bounds_proof}
	R \leq \dfrac{\ln{\big( 1-\frac{4}{3}\epsilon \big)} + \beta (\phi - F)}{\ln{\big( 1-\frac{4}{3}\epsilon' \big)} + \beta (\phi' - F')},
\end{equation}
where $F$ is the free energy of $\tau$,  and 
\begin{equation}
	\phi = -\beta^{-1} \log \zeta
\end{equation}
with $\zeta$ given by the expressions
\begin{align}
	\zeta &= \sum_{k\in \mathbb{Z}_3} \alpha_k e^{-\beta E_k}, \nonumber\\
	\alpha_k &= \<E_k| A_{\x_\star} |E_k\>,
\end{align}
and $W_\tau(\x)$ attains a minimum at $x=x_\star$. The primed variables are defined similarly for the output system.
\end{theorem}
\noindent The proof of this is provided in~\cref{thm:free-energy-bound-proof}, and is a generalization of the analysis for unital protocols.

>>>>

\begin{document}
% Labels and Refs (consistency, cref vs ref, thm vs lem vs prop, ??)
% Figs check(fig.0, move to app, consistency)

% Notation consistency (R is rate not bound, \ln{} -> \ln, italics)
% UK vs US english
% Spacing (figs, sections, paragraphs, eq. numbering, \noindent, QEDs)

@@

In this work, we develop a framework to analyse magic state distillation protocols where explicit physics of the system is incorporated.  Distillation bounds are obtained by considering how a given magic state protocol would behave if it were instead applied to a \emph{stabilizer} state that is distinguished by the physics of the system -- for example an equilibrium state at a characteristic temperature, or the maximally mixed state. The magic distillation bounds can then be expressed in terms of this distinguished degrees of freedom (e.g. free energy changes in the case of temperature, or non-unitality in the case of the maximally mixed state). Such bounds are therefore of interest in assessing how physical features such as temperature, noise-biases or fixed-point structure associated with restricted gate-sets constrain distillation protocols, and so are of relevance to existing work towards fault-tolerant quantum computing~\cite{Aliferis_2008, Stephens_2013, Li_2015, Babbush_2018, Tuckett_2019, Guillaud_2019, Fowler_2019}.

The approach we take is most closely aligned with resource theories of magic, however it differs in a few ways. Firstly, we obtain distillation upper bounds without the use of monotones, and secondly our distillation bounds are hardware dependent. We approach this problem using tools from statistical mechanics and recent work from the overlap of thermodynamics and quantum information theory~\cite{cit:janzing, cit:horodecki2, Brandao_2015, Vinjanampathy_2016, Goold_2016, cit:lostaglio, cit:gour}. In particular, we shall view \emph{magic as a form of free energy}, relative to the set of stabilizer states. From this perspective, the convex set of stabilizer states behaves as the set of ``equilibrium'' states, and magic corresponds to a form of non-equilibrium free energy that is strictly non-classical.

@@

Consider a $d$--dimensional quantum system with Hilbert space $\H_d$, and let $\{ |0\>, |1\>, \dots , |d-1\>\}$ denote the standard computational basis, defined over $\mathbb{Z}_d = \{ 0, 1, \dots,d-1 \}$. On this space, generalised Pauli matrices $X, Z$ can be defined by their respective roles as shift and phase operators, acting on the basis states as follows,
\begin{align}
    X \ket{k} &= \ket{k + 1} \label{eq:xpauli}\\
	Z \ket{k} &= \omega^k \ket{k}. \label{eq:zpauli}
\end{align}
Here $\omega \coloneqq e^{2\pi i/d}$ is the $d$-th root of unity and addition is taken modulo $d$. From these we can construct a phase space $\cal{P}_d = \mathbb{Z}_d \times \mathbb{Z}_d$ that provides a complete representation of the quantum system. Given a point $\bmx \coloneqq (x, p)$ we define a displacement operator, 
\begin{equation}\label{eq:ddef}
    D_{\bmx} \coloneqq \tau^{x p} X^{x} Z^{p},\ 
\end{equation}
where the phase factor $\tau \coloneqq -\omega^{1/2}$ ensures unitarity. We assume going forward that $d$ is an odd prime, however the case $d=2$ can also be handled, but with some additional technical caveats~\cite{Appleby_2005}. For a composite system with composite dimension $d = d_1 \dots d_n$ we can decompose the Hilbert space as $\cal{H}_d = \cal{H}_{d_1} \otimes \dots \otimes \cal{H}_{d_n}$, and then define displacement operators as
\begin{equation}\label{eq:composited}
    D_{\bmx} \coloneqq D_{(x_1, p_1)} \otimes \dots \otimes D_{(x_n, p_n)},
\end{equation}
where now we have
\begin{align*}
	\bmx \coloneqq (x_1, p_1, x_2, p_2, \dots, x_n, p_n) \in \cal{P}_{d_1} \times \dots \times \cal{P}_{d_n} \eqqcolon  \cal{P}_d,
\end{align*}
to denote the phase space point for the composite system. 
For simplicity going forward, we assume $n$ copies of a $d$--dimensional system $d_1=d_2 = \cdots = d$, and therefore we have that $\x \in \mathbb{Z}_d^{2n}$.

@@

\begin{figure}[t]
    \centering
    \subfigure[][]{%
    \label{fig:maxmix}%
    \includegraphics[height=2cm]{figs/maxmixed.pdf}
    }\hspace{8pt}%
    \subfigure[][]{%
    \label{fig:zero}%
    \includegraphics[height=2cm]{figs/zerostate.pdf}
    }\\
    \subfigure[][]{%
    \label{fig:bound}%
    \includegraphics[height=2cm]{figs/boundstate.pdf}
    }\hspace{8pt}%
    \subfigure[][]{%
    \label{fig:Strange}%
    \includegraphics[height=2cm]{figs/Strangestate.pdf}
    }
    \caption{\textbf{Qutrit Wigner distributions of varying magic.} 
    \subref{fig:maxmix} Maximally mixed state $\frac{1}{3}\id$; \subref{fig:zero} Stabilizer zero state $|0\>\<0|$; \subref{fig:bound} A non-stabiliser Wigner-positive state; \subref{fig:Strange} Magic Strange state $\ket{{\rm{S}}} = \frac{1}{\sqrt{2}}(\ket{1} - \ket{2})$, coined in~\cite{cit:veitch2}.
    }%
    \label{fig:wstate_examples}
\end{figure}

@@

Any magic state protocol will correspond to a quantum channel $\E$, and admits a Wigner representation $W_{\E}( \y | \x)$ that can be viewed as a transition matrix mapping phase space points $\x \rightarrow \y$. The representation obeys the relation 
\begin{equation}
	W_{\E(\rho)} (\y) = \sum_{\x \in \P_d} W_{\E}( \y | \x) W_\rho(\x),
\end{equation}
for any $\E$ and $\rho$. Since the magic protocol sends free states to free states, $\E$ sends all positively represented quantum state to other positively represented quantum states. In such cases, it can be shown~\cite{Wang_2019} that $W_{\E}(\y |\x)$ must form a stochastic matrix. In particular, all Clifford operations correspond to stochastic matrices in this Wigner representation.

@@

Our analysis applies majorization to magic states at the level of the associated Wigner distributions. However, these are in general quasi-probability distributions, so it is important to check how majorization is computed for these cases and what differences quasi-distributions bring over genuine probability distributions.

@@

However, if $\w$ is a \emph{quasi-probability} distribution with negative values, and $\r$ a regular probability distribution things are different. Now the Lorenz curve is no longer monotone increasing, but is a concave function that breaks through the $L_{\w|\r}(x) = 1$ barrier at an interior point and attains some non-trivial maximum $L_\star$ above the value $1$, before decreasing monotonically to $L_{\w|\r}(x)= 1$ at the end-point $x=1$. See~\cref{fig:lcs} for examples of non-monotonic Lorenz curves for quasi-distributions. Within quantum theory, the breaking of this barrier is associated with the degree of non-classicality. Since majorization provides a measure of order/disorder this is consistent with the intuition that negativity, for example within the context of quantum optics or quantum computing, can provide a greater form of order than is possible within classical theory.

\begin{figure}
    \centering
    \includegraphics[scale=0.35]{figs/lc_strange.pdf}
    \caption{\textbf{A `heretical' family of Lorenz curves.} Traditionally, Lorenz curves for distributions are monotone increasing cumulant functions that reach a maximum value of $1$. In contrast, Lorenz curves for magic states break through the value of $L(x)=1$ due to the presence of negativity in the associated quasi-probability distributions. The above family of curves correspond to multiple copies of noisy Strange states $\rho\coloneqq\rho_{\rm{S}}(\epsilon)^{\otimes n}$ (\cref{eq:noisysn}) for $n=2,4,6$ within a particular stabilizer fragment ($\sigma = \id/3$). Solid lines represent pure Strange states, while dashed lines represent $\epsilon$-noisy Strange states with depolarising error $\epsilon = 0.1$.
    }
    \label{fig:lcs}
\end{figure}

@@

In~\cref{sec:stab} we shall focus on the energetic costs of a protocol, that are defined due to the specific Hamiltonian that the system possesses. However, before that, in this section we encode physical constraints at a more abstract level in terms of the fixed-point structure of  the actual operations that are accessible to us. We show that any magic theory can always be decomposed in terms of `sub-theories' with a particular fixed-point structure. This section therefore exploits $\mathbf{d}$--majorization, while in~\cref{sec:stab} we drop the requirement of fixed-points and instead use the full relative majorization relation.

 We begin by defining the following sub-theory of any magic theory $\R$.
\begin{definition}\label{def:sigmafrag}
   Given a theory of magic $\R = (\F, \O)$, the \emph{$\sigma$--fragment of $\R$} is the sub-theory $\R_\sigma = (\F, \O_\sigma)$, in which the free operations are 
   \begin{equation}
        \O_\sigma \coloneqq \{ \E \in \O: \E(\sigma) = \sigma \},
    \end{equation}
namely those operations that leave the state $\sigma \in \F$ invariant.
\end{definition}
This notion provides a simple way to break up any theory into smaller, more manageable parts that are destinguished by additional physical constraints, illustrated in~\cref{fig:zoo}. Moreover, the union over all fragments returns the parent theory, so we are not discarding any information by breaking up the theory in this way. We make this statement precise as follows.
\begin{theorem}\label{thm:frag}
    Let $\R = (\F, \O)$ be a theory of magic.
Then a transformation $\rho \rightarrow \tau$ is possible in $\R$ if and only if the transformation $\rho \rightarrow \tau$ is possible in at least one $\sigma$--fragment of $\R$.
\end{theorem}
The proof of this is straightforward.
\begin{proof}
   Suppose the interconversion is possible in a $\sigma$--fragment via some $\E \in \O_\sigma$. But since $\O_\sigma \subseteq \O$ it is also possible in $\R$. Conversely, suppose the transformation is possible in $\R$ via some $\E$ in $\O$. The free states $\F$ are a closed, bounded set and moreover the image of $\F$ under the map $\E$ is in $\F$. By the Brouwer Fixed-Point theorem~\cite{cit:brouwer} this mapping must therefore have a fixed point $\sigma \in \F$. Therefore, $\E \in \O_\sigma$ and the interconversion is possible in this fragment.
\end{proof}

@@

If we consider some $\sigma \in \F$, and its corresponding fragment $\R_\sigma$, then we are typically interested in the ability to transform many copies of some magic state $\rho$ into more pure forms of magic. The state $\rho$ is assumed to have negativity in the Wigner representation, so $W_\rho(\x) < 0$ for some regions of $\x \in \P_d$. We again note that we restrict to odd-dimensional quantum systems, for simplicity. In contrast, the state $\sigma$ will have a Wigner distribution $W_\sigma(\x)$ that forms a genuine probability distribution. On the technical assumption that $\sigma$ is full rank, we also have that $W_\sigma(\x) > 0$ for all $\x \in \P_d$. Any non-full rank states $\sigma$ can be handled as a limiting case in which we inject an infinitesimal fraction of depolarising error $\epsilon (\id/d)$.

@@

If we have a magic state $\rho$ that has negativity in its Wigner representation, then, as discussed, its Lorenz curve over-shoots $1$ and reaches a non-trivial maximum $L_\star$ that depends on the particular state. There is an exact relation between $L_\star$ and sum-negativity or mana, which is provided by the following result. 
\begin{theorem}\label{lem:lcmax}
	Given a magic state $\rho$, the maximum $L_\star$ of its Lorenz curve $\lc{\rho}{\sigma}(x)$ is independent of the $\sigma$--fragment and equal to $1+\sn{\rho}$. Moreover, the majorization constraint is stronger than mana in every fragment.
\end{theorem}
The proof of this is given in~\cref{app:major}. Using the Lorenz curve perspective, it is simple to construct other monotones. For example, within a given fragment, the area above the horizontal line $y = 1$ is another magic monotone.
\begin{lemma}
Given a magic state $\rho$ and a free state $\sigma$, let $\A_\sigma(\rho)$ be the area of the region $\{(x, y): 1 \leq y \leq L_{\rho | \sigma}(x)\}$. Then $\A_\sigma(\rho)$ is a magic monotone for $\R_\sigma$.
\end{lemma}
\begin{proof}
Consider the transformation $\rho_1 \rightarrow \rho_2$ within $\R_\sigma$. We have that $L_{\rho_1|\sigma}(x)$ is never below the curve $L_{\rho_2|\sigma}(x)$, and therefore the region above $L=1$ for $\rho_2$ is a subset of the corresponding region for $\rho_1$. Thus, $\A_\sigma(\rho_1) \geq \A_\sigma(\rho_2)$, so $\A_\sigma$ is a magic monotone in $\R_\sigma$.
\end{proof}
In contrast to mana, the area monotone is \emph{specific to the fragment}, and its value will vary as we change $\sigma$. Therefore, its monotonicity is also including the physics of the fragment and provides a means to analyse magic distillation only under free operations that leave $\sigma$ invariant.

@@

For our distillation protocols we need to choose a canonical pure magic state as a target. For this we shall choose the Strange state $|S\>$ given as
\begin{equation}
|S\> \coloneqq \frac{1}{\sqrt{2}} (|1\> + |2\>),
\end{equation}
in the computational basis. This qutrit magic state is exceptionally symmetric and simple which benefits our analysis. Its distribution $W_S(\x)$ has a single negative value of $-1/3$ at $\x =(0,0)$ and the positive value $1/6$ at all other points. We can define the $\epsilon$--noisy Strange state as
\begin{equation}\label{eq:noisysn}
    \rho_{\rm{S}}(\epsilon)\coloneqq (1 - \epsilon) \ket{\rm{S}}\bra{\rm{S}} + \epsilon \frac{1}{3}\id,
\end{equation}
where $\epsilon$ is the depolarizing error parameter. Moreover, utilising the symmetries of the Strange state allows for any magic state $\rho$ to be processed via Clifford operations~\cite{cit:prakash,cit:prakash2} and put into this canonical form for some $\epsilon \ge 0$.

The Wigner distribution of the single-copy, $\epsilon$--noisy Strange state  is given by
\begin{equation}
	\W[\bmx]{\rho_{\rm{S}}(\epsilon)} = (1-\epsilon)\W[\bmx]{\rm{S}\>\<\rm{S}|} + \epsilon\W[\bmx]{\frac{1}{3}\id}.
\end{equation}
Since the Wigner distribution $W_{\I/3}(x)$ is the uniform probability distribution over the phase space, we get that $W_{\rho_S(\epsilon) } (\x)$ has a single negative component
\begin{equation}
	- v(\epsilon) \coloneqq - \left( \frac{1}{3} -\frac{4}{9}\epsilon \right)
\end{equation} 
located at the origin $\bmx = \bmo$ and positive components
\begin{equation}
	u(\epsilon) \coloneqq \frac{1}{6} -\frac{1}{18}\epsilon
\end{equation}
at the 8 phase space points $\x \ne (0,0)$. We assume that $\epsilon < 3/4$ to ensure the presence of negativity in the Wigner distribution.

@@

In the next section, we loosen the restriction of working in a fixed fragment, and instead use the full structure of relative majorization to obtain bounds that incorporate energetic aspects of a magic state protocol.

@@

\subsection{Free energy and sub-linear correlations in the thermodynamic limit}
To simplify our analysis we will make the following physical assumption. We assume that in the asymptotic/thermodynamic limit $n,m \rightarrow \infty$ that correlations generated between subsystems are negligible. This implies that $d\mu(x)$ is peaked on a particular state $\tau'$, so $\E(\tau) = \tau^{\prime \otimes m}$. Physically, this is a natural scenario -- for example, in the context of traditional thermodynamics, it states that the output system is well-described by intrinsic variables that are encoded in $\tau'$, and correlations are sub-linear in $m$. This assumption allows us to associate a free energy per qudit of the output state, which would be more complex if non-trivial correlations were present.

@@

\subsection{Temperature-dependent distillation bounds}
The preceding discussion gives a means to specify energetic details of a protocol through how much it would disturb some reference equilibrium state. When combined with relative majorization, it also allows us to derive magic distillation bounds that combine both the computational measures $\epsilon,\epsilon'$, with terms that depend on the Hamiltonian of the physical system. As discussed in section~\nick{REF}, from the perspective of majorization the presence of magic can be viewed as a form of non-classical free energy, so it is also of interest to make an explicit link between the amount of magic in a non-equilibrium state $\rho_S(\epsilon)$ and the actual free energy of an equilibrium state $\tau(\beta)$.

@@

Then the magic distillation rate $R = m/n$ for the protocol is bounded by the expression
\begin{equation}\label{eq:rate_bounds_proof}
	R \leq \dfrac{\ln{\big( 1-\frac{4}{3}\epsilon \big)} + \beta (\phi - F)}{\ln{\big( 1-\frac{4}{3}\epsilon' \big)} + \beta (\phi' - F')},
\end{equation}
where $F$ is the free energy of $\tau$,  and 
\begin{equation}
	\phi = -\beta^{-1} \log \zeta
\end{equation}
with $\zeta$ given by the expressions
\begin{align}
	\zeta &= \sum_{k\in \mathbb{Z}_3} \alpha_k e^{-\beta E_k}, \nonumber\\
	\alpha_k &= \<E_k| A_{\x_\star} |E_k\>,
\end{align}
and $W_\tau(\x)$ attaining a minimum at $\x=\x_\star$. The primed variables are defined similarly for the output system.
\end{theorem}
\noindent The proof of this is provided in~\cref{thm:free-energy-bound-proof}, and is a generalization of the analysis for unital protocols.

The above bound depends on: (a) quantum computational measures $\epsilon, \epsilon'$, (b) thermodynamic quantities $F, F',$  and (c) intermediate terms $\phi, \phi'$. The intermediate terms depend on how the energy eigenbasis of the system $\{\ket{E_k}\}$ relates to the computational stabilizer basis $\{\ket{k}\}$.  In particular, the energetic term $\phi$ quantifies the degree to which the negativity in the Wigner representation can be ascribed a sharp energetic value for the given Hamiltonian of the system. Its form is similar to $F$, and it can informally be viewed as a `magic free energy' term, associated to what can informally be viewed as a `partition function' $\zeta$. The coefficients $\alpha_k$ may be negative for some $k$, when the Hamiltonian eigen-decomposes into non-stabilizer states, but the quantity $\phi$ is always well-defined, since the function $\zeta$ is a positive multiple of the stabilizer Wigner component $W_\tau(0,0)$, as seen in the proof of the theorem. The quantity $\phi$ would diverge if the state $\tau$ acquired zero Wigner components, which only occurs if $\tau$ is on the boundary of the set of Wigner-positive states.

In the case that the system's Hamiltonian $H$ has an eigenbasis consisting of stabilizer states then we show in~\cref{app:main_proof} that $\phi$ takes on a sharp, temperature-independent energy value $\phi \in \{E_0,E_1,E_2\}$.
